Low-Cost Data Deduplication for Virtual Machine Backup in Cloud Storage

### 2013-05-14 hotstorage13/arch.tex
save some lines

\item {\bf Separation of duplicate detection and data backup.}
%Request accumulation and partition-based deduplication.}
The second level detection requires a global comparison of fingerprints.
% of stored chunks with the new chunks scanned during the backup process. 
%Bloom filters and caching allows some of content 
%fingerprints to be stored on the cheap disks~\cite{bottleneck08}. Such an optimization is good 
%for inline deduplication where chunks scanned need to be determined to be duplicate or not instantly without 
%waiting while  it memory consumption is still significant in competing with stand virtual 
%machine activities.  
Our approach is to perform duplicate detection first before actual data backup.
That requires a prescanning of  data  blocks in dirty VM segments, which
%  and after that, real backup  for those non-duplicates is conducted. 
does incur an extra  round of VM reading.
% while avoiding inline deduplication and leading to a much smaller resource requirement.  
During VM prescanning, detection requests are accumulated.
Aggregated duplicate requests can be processed partition by partition. 
%We also accumulate the delete requests and perform them partition by partitions. 
Since each partition corresponds to a small portion of global content index, 
memory cost to process detection requests within a partition is small.
%the cost to process detection requests within a partition is significantly smaller than processing all 
%the requests globally.

@@

%\item 
Stage 3 is to perform real backup one VM at a time.
The system loads the duplicate summary of a VM, 
reads  dirty segments of a VM, and outputs non-duplicate blocks to the final backup 
storage. Additionally, the global index on each machine is updated with the meta data of new chunk blocks. 
When a segment is not dirty, the system only needs to output the segment meta data such as a reference pointer. 
There is an option to directly read dirty blocks instead of fetching a dirty segment which can include duplicate
blocks. Our experiment shows that it is faster to read dirty segments in the tested workload.
Another issue is that during global index update after new block creation in the backup storage,
the system  may find some  blocks with the same fingerprints have been 
created redundantly. For example, two different VM blocks that have the same  fingerprint are not detected
because  the  global index has not contained such a fingerprint yet. 
The redundancy is discovered and logged during the index update and can be repaired
periodically when necessary.  Our experience is that there is a redundancy during the initial snapshot backup and once 
that is repaired, the percentage of redundant blocks due to concurrent processing  is insignificant.

>>>>

\item {\bf Separation of duplicate detection and data backup.}
%Request accumulation and partition-based deduplication.}
The second level detection requires a global comparison of fingerprints.
% of stored chunks with the new chunks scanned during the backup process. 
%Bloom filters and caching allows some of content 
%fingerprints to be stored on the cheap disks~\cite{bottleneck08}. Such an optimization is good 
%for inline deduplication where chunks scanned need to be determined to be duplicate or not instantly without 
%waiting while  it memory consumption is still significant in competing with stand virtual 
%machine activities.  
Our approach is to perform duplicate detection first before actual data backup.
That requires a prescanning of  data  blocks in dirty VM segments, which
%  and after that, real backup  for those non-duplicates is conducted. 
does incur an extra  round of VM reading.
% while avoiding inline deduplication and leading to a much smaller resource requirement.  
During VM prescanning, detection requests are accumulated.
Aggregated deduplicate requests can be processed partition by partition. 
%We also accumulate the delete requests and perform them partition by partitions. 
Since each partition corresponds to a small portion of global index, 
memory cost to process detection requests within a partition is small.
%the cost to process detection requests within a partition is significantly smaller than processing all 
%the requests globally.

@@

%\item 
Stage 3 is to perform real backup one VM at a time.
The system loads duplicate summary of a VM, 
reads  dirty segments of a VM, and outputs non-duplicate blocks to the final backup 
storage. Additionally, the global index on each machine is updated with the meta data of new chunk blocks. 
When a segment is not dirty, the system only needs to output the segment meta data such as a reference pointer. 
There is an option to directly read dirty blocks instead of fetching a dirty segment which can include duplicate
blocks. Our experiment shows that it is faster to read dirty segments in the tested workload.
Another issue is that during global index update after new block creation,
% in the backup storage,
the system  may find some  blocks with the same fingerprints have been 
created redundantly. For example, two different VM blocks that have the same  fingerprint are not detected
because  the  global index has not contained such a fingerprint yet. 
The redundancy is discovered and logged during the index update and can be repaired
periodically when necessary.  Our experience is that there is a redundancy during the initial snapshot backup and once 
that is repaired, the percentage of redundant blocks due to concurrent processing  is insignificant.

### 2013-05-14 hotstorage13/background.tex
save some lines

%The previous work for storage backup has extensively used  data deduplication techniques can eliminate redundancy globally among different files from different users.
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.
Offline deduplication is used in ~\cite{EMC,NetAppOffline} to remove previously written duplicate blocks during idle time.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data blocks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
%Still such methods use a significant amount of memory resource for filtering and caching.
Such methods use dedicated nodes for deduplication, which add cost to cloud infrastructure and is less scalable.
%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%Deepavali et al.~\cite{extreme_binning09} and Zhang et al.~\cite{WeiZhangIEEE}  
to reduce memory requirement with a tradeoff of the reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup  
In comparison, this paper focuses on  full deduplication without approximation.

>>>>

%The previous work for storage backup has extensively used  data deduplication techniques can eliminate redundancy globally among different files from different users.
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.
Offline deduplication is used in ~\cite{EMC,NetAppOffline} to remove previously written duplicate blocks during idle time.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data blocks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
%Still such methods use a significant amount of memory resource for filtering and caching.
%Such methods use dedicated nodes for deduplication, which add cost to cloud infrastructure and are less scalable.
As discussed in Section~\ref{sect:intro},
there is no dedicated resource for deduplication in our targeted setting and low memory usage
 is required so that the resource impact to other cloud services is minimized.
%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%Deepavali et al.~\cite{extreme_binning09} and Zhang et al.~\cite{WeiZhangIEEE}  
to reduce memory requirement with a tradeoff of the reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup  
In comparison, this paper focuses on  full deduplication without approximation.

### 2013-05-14 hotstorage13/intro.tex
re
save some lines

\section{Introduction}

%Virtualization is the engine behind many popular cloud computing platforms.
%Amazon, Alibaba,  and many others have provided public VM clouds that host
%tens of thousands of  virtual machines(VM) created for various active users.
Periodic  archiving of virtual machine (VM) snapshots is important 
for long-term data  retention and fault recovery. 
%in case of machine failures.  
For example, daily backup of VM images  is conducted automatically 
at Alibaba which provides the largest public cloud service in China.
The cost of frequent backup of VM snapshots is  high because of the huge storage demand.
This issue has been addressed by   storage data deduplication ~\cite{venti02,bottleneck08} that
identifies redundant content duplicates among snapshots.  
One architectural approach is to attach  a separate backup system with deduplication
support to the cloud cluster, and  every machine periodically transfers snapshots to
the attached backup system.  The cost of allocating such a dedicated backup  system can be expensive.

This paper seeks for a low-cost architecture option and considers that
a backup service uses the existing cloud computing resource.
%Since most of backup data is not used in practice, system resource in such a service is not fully utilized. 
Performing deduplication adds significant  memory cost for comparison of content fingerprints. 
Since each physical machine in a cluster  hosts many VMs, memory contention happens frequently. 
Cloud providers often wish that the backup service only consumes  small or modest resources 
with a minimal impact to the existing cloud services.  Another challenge for backup with deduplication is 
that deletion of old snapshots compete for computing resource as well. That is because data dependence created 
by duplicate relationship among snapshots  adds processing complexity.
% especially when  VMs can migrate around in the cloud.

>>>>

\section{Introduction}
\label{sect:intro}

%Virtualization is the engine behind many popular cloud computing platforms.
%Amazon, Alibaba,  and many others have provided public VM clouds that host
%tens of thousands of  virtual machines(VM) created for various active users.
Periodic  archiving of virtual machine (VM) snapshots is important 
for long-term data  retention and fault recovery. 
%in case of machine failures.  
For example, daily backup of VM images  is conducted automatically 
at Alibaba which provides the largest public cloud service in China.
The cost of frequent backup of VM snapshots is  high because of the huge storage demand.
This issue has been addressed by   storage data deduplication ~\cite{venti02,bottleneck08} that
identifies redundant content duplicates among snapshots.  
One architectural approach is to attach  a separate backup system with deduplication
support to the cloud cluster, and  every machine periodically transfers snapshots to
the attached backup system.  Such a dedicated backup  configuration can be expensive, 
considering that significant networking  and computing resource  is required to transfer 
raw data and conduct signature comparison.

This paper seeks for a low-cost architecture option and considers that
a backup service uses the existing cloud computing resource.
%Since most of backup data is not used in practice, system resource in such a service is not fully utilized. 
Performing deduplication adds significant  memory cost for comparison of content fingerprints. 
Since each physical machine in a cluster  hosts many VMs, memory contention happens frequently. 
Cloud providers often wish that the backup service only consumes  small or modest resources 
%with a minimal impact to the existing cloud services.  Another challenge for backup with deduplication is 
with a minimal impact to the existing cloud services.  Another challenge is 
%that deletion of old snapshots compete for computing resource as well. That is because data dependence created 
that deletion of old snapshots compete for computing resource as well, because data dependence created 
by duplicate relationship among snapshots  adds processing complexity.
% especially when  VMs can migrate around in the cloud.
